Parameter Estimation for the Compound Poisson Model

### 2017-01-09 reports/compoundPoisson/compoundPoisson.tex
compound Poisson proof read

\title{Parameter Estimation for the Compound Poisson Model}
\author{Sherman Ip \quad \today}

@@

\section{Introduction}
In the mini-project, a compound Poisson model was used to model the distribution of greyvalues in a pixel. The model was built by assuming that photon arrivals are a Poisson process and that each photon has random energy, distributed according to the energy spectrum. The greyvalue is then proportional to the sum of energies of each photon detected in some exposure time, with some Normal noise added.

@@

For a given i.i.d.~random sample of compound Poisson-Gamma distributed data, the unknown parameters $\nu$, $\alpha$ and $\lambda$, can be estimated using method of moments. Suppose $\widehat{\mu}$ is an estimator of $\expectation[X]$ and $\widehat{\mu}_j$ is an estimator of $\expectation\left[\left(X-\mu\right)^j\right]$ for $j=2,3$. Then the estimators
\begin{equation}
\widehat{\nu}=\frac{\widehat{\mu}^2\widehat{\mu}_2}{\tau\left(2\widehat{\mu}_2^2-\widehat{\mu}_3\widehat{\mu}\right)}
\end{equation}
\begin{equation}
\widehat{\alpha}=\frac{\widehat{\mu}_3\widehat{\mu}-2\widehat{\mu}_2^2}{\widehat{\mu}_2^2-\widehat{\mu}\widehat{\mu}_3}
\end{equation}
\begin{equation}
\widehat{\lambda}=\frac{\widehat{\mu}\widehat{\mu}_2}{\widehat{\mu}\widehat{\mu}_3-\widehat{\mu}_2^2}
\end{equation}
are method of moments estimators of $\nu$, $\alpha$ and $\lambda$ respectively \cite{withers2011compound}. These estimators suffer because estimation is not done through the sufficient statistics and can be negative, this is a problem because the parameters do not take non-positive values.

@@

\section{Future work}
The \texttt{R} package \texttt{cplm} \cite{zhang2013likelihood} numerically estimate parameters for the compound Poisson in the GLM framework. However for large parameters, there were numerical problems and the estimators' performance were hit and miss. Future work will reply on this package and the saddlepoint approximation to estimate parameters for the compound Poisson. Numerical problems can come from ill-conditioned variables and variables need to be scaled. One useful property of the compound Poisson is that the parameter $\lambda$ can absorb any scaling by a constant. That is if one can express a compound Poisson random variable as
\begin{equation}
X\sim\text{CompoundPoisson}\left(\nu\tau,\alpha,\lambda\right)
\end{equation}
then for a positive constant $k$
\begin{equation}
kX\sim\text{CompoundPoisson}\left(\nu\tau,\alpha,\lambda/k\right) \ .
\end{equation}

>>>>

\title{Parameter Estimation for the Compound Poisson Model}
\author{Sherman Ip \quad January 9, 2017}

@@

\section{Introduction}
In the mini-project, a compound Poisson model was used to model the distribution of greyvalues in a pixel. The model was built by assuming that photon arrivals are a Poisson process and that each photon has random energy, distributed according to the energy spectrum. The greyvalue was then proportional to the sum of energies of each photon detected in some exposure time, with some Normal noise added.

@@

For a given i.i.d.~random sample of a compound Poisson-Gamma random variable, the unknown parameters $\nu$, $\alpha$ and $\lambda$, can be estimated using method of moments. Suppose $\widehat{\mu}$ is an estimator of $\expectation[X]$ and $\widehat{\mu}_j$ is an estimator of $\expectation\left[\left(X-\mu\right)^j\right]$ for $j=2,3$. Then the estimators
\begin{equation}
\widehat{\nu}=\frac{\widehat{\mu}^2\widehat{\mu}_2}{\tau\left(2\widehat{\mu}_2^2-\widehat{\mu}_3\widehat{\mu}\right)}
\end{equation}
\begin{equation}
\widehat{\alpha}=\frac{\widehat{\mu}_3\widehat{\mu}-2\widehat{\mu}_2^2}{\widehat{\mu}_2^2-\widehat{\mu}\widehat{\mu}_3}
\end{equation}
\begin{equation}
\widehat{\lambda}=\frac{\widehat{\mu}\widehat{\mu}_2}{\widehat{\mu}\widehat{\mu}_3-\widehat{\mu}_2^2}
\end{equation}
are method of moments estimators of $\nu$, $\alpha$ and $\lambda$ respectively \cite{withers2011compound}. These estimators suffer because estimation is not done through the sufficient statistics and can be negative, this is a problem because the parameters do not take non-positive values.

@@

\section{Future work}
The \texttt{R} package \texttt{cplm} \cite{zhang2013likelihood} numerically estimate parameters for the compound Poisson in the GLM framework. However for large parameters, there were numerical problems and the estimators' performance were hit and miss. Future work will rely on this package and the saddlepoint approximation to estimate parameters for the compound Poisson. Numerical problems can come from ill-conditioned variables and variables need to be scaled. One useful property of the compound Poisson is that the parameter $\lambda$ can absorb any scaling by a constant. That is if one can express a compound Poisson random variable as
\begin{equation}
X\sim\text{CompoundPoisson}\left(\nu\tau,\alpha,\lambda\right)
\end{equation}
then for a positive constant $k$
\begin{equation}
kX\sim\text{CompoundPoisson}\left(\nu\tau,\alpha,\lambda/k\right) \ .
\end{equation}
